---
title: "Final Project"
author: "Ian Baker, Loughlin Claus, Zack Schieberl"
date: "12/7/2019"
geometry: margin=2cm
output: pdf_document
---

## Pledge
I pledge my honor that I have abided by the Stevens Honor System
- Ian Baker, Loughlin Claus, Zack Schieberl

## 11.53

```{r fig.height=3, fig.width=3.5, fig.show='hold'}
cheese <- as.matrix(read.csv2("cheese.csv", header = TRUE, sep = ","))

printRegEq <- function(funcSum, name) {
  cat("Taste vs", name, funcSum$coefficients[, 3][1], "+", name, "*", funcSum$coefficients[, 3][2], "\n") 
}

cheeseCols <- colnames(cheese)
for (col in cheeseCols) {
  cur <- as.numeric(cheese[, col])
  # mean, median, sd, iqr
  out <- c(paste("Type:", col), paste("Mean:", round(mean(cur), 2)),
           paste("Median:", round(median(cur), 2)), paste("SD:", round(sd(cur), 2)),
           paste("IQR:", round(IQR(cur), 2)))
  print(format(out, justify = "left", trim = TRUE))
  # stemplot
  stem(cur)
  # normal quantile plot
  qqnorm(cur, main = col)
  qqline(cur)
}
```

While H2S and Taste have some right skew, and Acetic has two peaks, the data all appears to be relatively normal. There are no outliers in the data.

## 11.54

```{r fig.height=3, fig.width=3.5, fig.show='hold'}
for (col in cheeseCols) {
  colIdx <- grep(col, cheeseCols)
  col1Data <- as.numeric(cheese[, col])
  for (col2 in cheeseCols) {
    if (colIdx < grep(col2, cheeseCols)) {
      col2Data <- as.numeric(cheese[, col2])
      plot(col1Data, col2Data, xlab = col, ylab = col2)
      correl <- cor.test(col1Data, col2Data)
      cat("Correlation between", col, "and", col2, "is:", correl$estimate,
          "with a p-value of", correl$p.value, "\n")
    }
  }
}
```

## 11.55

```{r}
tasteCol <- as.numeric(cheese[, "taste"])
aceticCol <- as.numeric(cheese[, "acetic"])
tasteVsAcetic <- lm(tasteCol ~ aceticCol, data.frame(cheese))
plot(aceticCol, tasteCol, xlab = "Acetic", ylab = "Taste")
abline(tasteVsAcetic)
```

```{r fig.height=3, fig.width=3.5, fig.show='hold'}
tVsAResiduals <- residuals(tasteVsAcetic)
plot(cheese[, "h2s"], tVsAResiduals, xlab = "H2S", ylab = "TvA Residuals")
plot(cheese[, "lactic"], tVsAResiduals, xlab = "Lactic", ylab = "TvA Residuals")

printRegEq(summary(tasteVsAcetic), "Acetic")
plot(density(tVsAResiduals))
```

The residuals both have a normal distribution and seem to be positively associated with Lactic and H2S.

## 11.56

```{r}
h2sCol <- as.numeric(cheese[, "h2s"])
tasteVsH2S <- lm(tasteCol ~ h2sCol, data.frame(cheese))
plot(h2sCol, tasteCol, xlab = "H2S", ylab = "Taste")
abline(tasteVsH2S)
```

```{r fig.height=3, fig.width=3.5, fig.show='hold'}
tVsHResiduals <- residuals(tasteVsH2S)
plot(cheese[, "acetic"], tVsHResiduals, xlab = "Acetic", ylab = "TvH2S Residuals")
plot(cheese[, "lactic"], tVsHResiduals, xlab = "Lactic", ylab = "TvH2S Residuals")

printRegEq(summary(tasteVsH2S), "H2S")
plot(density(tVsHResiduals))
```

The residuals appear to have a normal distribution. From the graphs there appears to be no correlation between the residuals and other variables. 

## 11.57

```{r}
lacticCol <- as.numeric(cheese[, "lactic"])
tasteVsLactic <- lm(tasteCol ~ lacticCol, data.frame(cheese))
plot(lacticCol, tasteCol, xlab = "Lactic", ylab = "Taste")
abline(tasteVsLactic)
```

```{r fig.height=3, fig.width=3.5, fig.show='hold'}
tVsLResiduals <- residuals(tasteVsLactic)
plot(cheese[, "acetic"], tVsLResiduals, xlab = "Acetic", ylab = "TvL Residuals")
plot(cheese[, "h2s"], tVsLResiduals, xlab = "H2S", ylab = "TvL Residuals")

printRegEq(summary(tasteVsLactic), "Lactic")
plot(density(tVsLResiduals))
```

Again, the residuals appear to have a normal distribution and there appears to be no correlation between the residuals and the other variables.

## 11.58

```{r results="asis"}
tVsASum <- summary(tasteVsAcetic)
tVsHSum <- summary(tasteVsH2S)
tVsLSum <- summary(tasteVsLactic)

fStats  <- c(tVsASum$fstatistic[1], tVsHSum$fstatistic[1], tVsLSum$fstatistic[1])
pVals   <- c(tVsASum$coefficients[, 4][2], tVsHSum$coefficients[, 4][2],
             tVsLSum$coefficients[, 4][2])
rSqVals <- c(tVsASum$r.squared, tVsHSum$r.squared, tVsLSum$r.squared)
sdEst   <- c(tVsASum$sigma, tVsHSum$sigma, tVsLSum$sigma)

knitr::kable(data.frame(
  fStats, pVals, rSqVals, sdEst
))
```

```{r}
printRegEq(tVsASum, "Acetic")
printRegEq(tVsHSum, "H2S")
printRegEq(tVsLSum, "Lactic")
```

The intercepts in the three equations are different because the explanitory variables all have different values, leading to the points being plotted in different places. Since the linear equations are estimating the best fit line for these datapoints, it is only natural that the differing data produces different intercepts.

## 11.59

```{r}
tasteVsAnH <- lm(tasteCol ~ aceticCol+h2sCol, data.frame(cheese))
tvAHsum<-summary(tasteVsAnH)
TAHcoef<-unlist(tasteVsAnH["coefficients"])
cat("Taste^ =", TAHcoef[1],"+",TAHcoef[2],"* acetic +",TAHcoef[3],"* H2S\n")
cat("p-value:",1-pf(tvAHsum[[10]][1],tvAHsum[[10]][2],tvAHsum[[10]][3]),'\n')
coef(tvAHsum)
```

There is not much statistical significance of Acetic in this model (seen in Acetic's large p-value of 0.406), this loss in (Acetic's) significance from its prior model (which used Acetic as the sole predictor) is most likely due to the large positive correlation between H2S and Acetic. When Acetic is used to predict Taste in conjunction with H2S, there is not much significant information contributed by Acetic that H2S hasn't already done better. This can be seen in the realtively small amount of change Acetic causes between the p-values of this model (7.645e-6) and the p-values of H2S' model (1.374e-6).

Between the model using both Acetic and H2S as a predictor and the model using solely Acetic as the predictor, I would prefer the model using both Acetic and H2S as a predictor, because of the model's much smaller p-value (7.645e-6 vs 0.00168). Although, if I had the choice, I would pick H2S over both of them, as it has an even smaller p-value (1.374e-6).


## 11.60

```{r}
THLModel <- lm(tasteCol ~ (h2sCol + lacticCol), data.frame(cheese))
THLSum   <- summary(THLModel)

cat("Regression Equation for H2S and Lactic:",
    THLSum$coefficients[, 3][1], "+", "H2s+Lactic *", THLSum$coefficients[, 3][2], "\n")

lmp <- function(fStats) {
  return (pf(fStats[1], fStats[2], fStats[3], lower.tail = F))
}

cat("P-Value:", lmp(THLSum$fstatistic), "\n")
```

Since the p-value of the model with both variables is far less than the p-values of the variables by themselves, it is a much better fit for predicting cheese taste.

## 11.61

```{r}
tasteVall <- lm(tasteCol ~ (aceticCol + h2sCol + lacticCol), data.frame(cheese))
allSum<-summary(tasteVall)
tasteVallcoef <-unlist(tasteVall["coefficients"])
words<-colnames(cheese)
cat("Taste^ =", tasteVallcoef[1])
for (i in c(2,3,4)){
  cat(" +", tasteVallcoef[i],"*",words[i])
  plot(cheese[,i],residuals(tasteVall),xlab=words[i],ylab="Residuals",
       main=paste("Scatter plot of",words[i],"and all Residuals"))
}
hist(residuals(tasteVall),xlab="all Residuals", main="Histogram of all Residuals")
cat('\np-value:',1-pf(allSum[[10]][1],allSum[[10]][2],allSum[[10]][3]),'\n')
coef(allSum)
boxplot(residuals(tasteVall),main="Boxplot of all Residuals")
```
The Acetic explanatory variable is practically negligable (has a p-value of practically 1, and its coefficient in the linear regression equation is practically 0), which leaves H2S and Lactice as the only statistically significant explanatory variables in this model.
None of the scatter plots of the Residuals vs the individual explanatory variables seem to display any notable patterns. The Residuals' histogram and Boxplots seem to display a normally distribution, though there is an outlier with the value of 25.44886. 
Of all the Models computed, I would most prefer to use the model that uses h2s and lactic to predict taste, as not only are they the only statistically significant variables in this problem, but their individual model also has the smallest p-value (6.551e-07) of all the models (meaning it is the most statisticallly significant of all the models). 